---
title: "EpilepsyTreatmentAnalysis"
author: "Duc-Anh Nguyen"
date: "2025-02-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


The dataset epil contains information of treatments of 59 patients of epilepsy. 
Goal: analysis the variable y: number of seizures (caused by epilepsy), measured in 4 periods, each are 2 weeks long.
There are 59 patients (covariate subject) and the other covariates are:  
+ trt Behandlung(placebo or probagide)
+ lbase: log-transformed and centered of number of seizures the whole study (8 weeks)
+ lage: log-transformed and centered of age
+ V4 indicator of the fourth period 
+ period 

As we are modeling count data we are thinking of poisson linear model
```{r}
library(MASS)
library(tidyverse)
data("epil")
head(epil)

skimr::skim(epil)
table(epil$y)
table(epil$trt)
table(epil$lbase)
table(epil$lage)
table(epil$V4)
table(epil$period)

ggplot(aes(period, y, col = as.factor(subject)), data = epil) + geom_point() + geom_line() +facet_wrap(~trt)



```


```{r}
library(lme4)
mixed_poisson_model <- glmer(
  formula = y ~ lbase*trt + lage + V4 +(1|subject),
  family = poisson,
  data = epil
)
summary(mixed_poisson_model)


```


```{r}
# install.packages("gee")
library(gee)

# for correlation structure "independence
model_gee_indep <-
  gee(
    formula = y ~ lbase * trt + lage + V4,
    id = subject,
    family = poisson,
    corstr = "independence",
    data = epil
  )

# correlation structure "exchangeable"
model_gee_equi <- gee(
  formula = y ~ lbase * trt + lage + V4,
  id = subject,
  family = poisson,
  corstr = "exchangeable",
  data = epil
)


# correlation structure "AR-1"
model_gee_AR1 <- gee(
  formula = y ~ lbase * trt + lage + V4,
  id = subject,
  family = poisson,
  corstr = "AR-M",
  Mv = 1,
  data = epil
)

model_Poisson <- glm(
  formula = y ~ lbase + trt + lage + V4 + lbase:trt,
  family = poisson,
  data = epil
)
model_quasi_Poisson <- glm(
  formula = y ~ lbase + trt + lage + V4 + lbase:trt,
  family = quasipoisson,
  data = epil
)



beta_hat <- se_beta_hat <- as.data.frame(matrix(0, 6, 6))
colnames(beta_hat) <-
  colnames(se_beta_hat) <- names(model_Poisson$coefficients)
rownames(beta_hat) <-
  rownames(se_beta_hat) <- c("Poisson",
                             "Quasi-Poisson",
                             "GEEIndep",
                             "GEEEqui",
                             "GEEAR-1",
                             "RandomIntercept")
beta_hat[1, ] <- model_Poisson$coefficients
beta_hat[2, ] <- model_quasi_Poisson$coefficients
beta_hat[3, ] <- model_gee_indep$coefficients
beta_hat[4, ] <- model_gee_equi$coefficients
beta_hat[5, ] <- model_gee_AR1$coefficients
beta_hat[6, ] <- mixed_poisson_model@beta
se_beta_hat[1, ] <- sqrt(diag(summary(model_Poisson)$cov.unscaled))
se_beta_hat[2, ] <- sqrt(diag(summary(model_quasi_Poisson)$cov.scaled))
se_beta_hat[3, ] <- sqrt(diag(model_gee_indep$robust.variance))
se_beta_hat[4, ] <- sqrt(diag(model_gee_equi$robust.variance))
se_beta_hat[5, ] <- sqrt(diag(model_gee_AR1$robust.variance))
se_beta_hat[6, ] <- sqrt(diag(summary(mixed_poisson_model)$vcov))
beta_hat
se_beta_hat


summary(model_gee_indep)

summary(model_gee_equi)

summary(model_gee_AR1)
```


The estimators:
+ pretty much the same for the mixed lm of poisson, quasi poisson and marginal one


The se:
+ the differences between the mixed lm and the marginal(except for V4) are quite significant
 